Large and Infinitary Quotient Inductive-Inductive Types

### 2020-04-11 draft/paper.tex
fixes

Cartmell \cite{gat} defines generalized algebraic theories (GATs) using type
theoretic syntax. Compared to our QII signatures, his signatures support sort
equations but do not cover infinitary constructors or recursive equations. A way
to encode sort equations in our signatures is using isomorphisms instead of
equalities. In contrast to our algebraic definition, Cartmell's signatures are
given by presyntax, named variables and typing relations, there is no explicit
model theory provided for signatures, and no explicit term model construction is
given. Cartmell focuses instead on showing that contextual categories serve as
classifying categories for GATs.

A more semantic approach to QIITs is given by Altenkirch et al.\
\cite{gabe}. They generalize the initial algebra semantics of
inductive types to QIITs by considering towers of functors and
building complete categories of algebras from such towers. Their
notion of signature does not enforce strict positivity, hence
describes a larger class of QII signatures. The equivalence of
initiality and induction is showed, but initial algebras are not
constructed.

@@

Higher inductive types (HITs) are generalizations of QIITs in settings with
proof-relevant identity types. They were introduced before QIITs
\cite{HoTTbook}. \cite{hitsinprogramming} is a description of a subset of HITs. \cite{hiit}
describes a syntax for higher inductive-inductive types using a theory of
signatures similar to ours, however does not construct categories of algebras or
show initiality. Semantics for different subclasses of HITs are given by
\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}. Cubical
type theories were shown to support HITs in a computational way
\cite{niels,hiit}.

@@

With regards to the reduction of QIITs to simple type formers, the reduction of
infinitary QIITs appears to be more challenging than the finitary
case. \cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
constructible from inductive types and simple quotients with relations. In the
finitary case, a generalization of the approach in \cite{induction-is-enough}
seems promising ; this amounts to a Streicher-style initial algebra construction
\cite{streicher2012semantics} for the theory of finitary QIIT signatures. In
particular, Brunerie et al. \cite{brunerie} have formalized in Agda this
construction for a comparable type theory, using UIP, function extensionality,
propositional extensionality and simple quotient types.

>>>>

Cartmell \cite{gat} defines generalized algebraic theories (GATs) using
type-theoretic syntax. Compared to our QII signatures, his signatures support
sort equations but do not cover infinitary constructors or recursive
equations. A way to encode sort equations in our signatures is using
isomorphisms instead of equalities. In contrast to our algebraic definition,
Cartmell's signatures are given by presyntax, named variables and typing
relations, there is no explicit model theory provided for signatures, and no
explicit term model construction is given. Cartmell focuses instead on showing
that contextual categories serve as classifying categories for GATs.

A more semantic approach to QIITs is given by Altenkirch et
al.\ \cite{qiits}. They generalize the initial algebra semantics of inductive
types to QIITs by considering towers of functors and building complete
categories of algebras from them. Their notion of signature does not enforce
strict positivity, hence describes a larger class of QII signatures. They show
equivalence of initiality and induction, but do not construct initial algebras,
nor would that be possible because of the lack of a positivity restriction.

@@

Higher inductive types (HITs) are generalizations of QIITs in settings with
proof-relevant identity types. They were introduced before QIITs
\cite{HoTTbook}. \cite{hiit} describes a syntax for higher inductive-inductive
types using a theory of signatures similar to ours, but it does not construct
categories of algebras and initial algebras. Semantics for different subclasses
of HITs are given by
\cite{lumsdaineShulman,10.1145/3209108.3209130,nielsmsc,sojakova,moeneclaey}. Cubical
type theories were shown to support some HITs in a computational way
\cite{cubicalhits,10.1145/3290314}.

@@

With regards to the reduction of QIITs to simple type formers, the reduction of
infinitary QIITs appears to be more challenging than the finitary
case. \cite[Section 9]{lumsdaineShulman} shows that infinitary QIITs are not
constructible from inductive types and simple quotients with relations. In the
finitary case, a generalization of the approach in \cite{induction-is-enough}
seems promising; this amounts to a Streicher-style initial algebra construction
\cite{streicher2012semantics} for the theory of finitary QIIT signatures. In
particular, Brunerie et al. \cite{brunerie} have formalized in Agda this
construction for a comparable type theory, using UIP, function extensionality,
propositional extensionality and simple quotient types.
